Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-18 src/tex/ms.tex
Update

In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO)
will finish its maintenance and start observing new gravitational wave (GW)
results. This new O4 run is expected to double the rate of current binary black
hole (BBH) observations \citep{abbott2020prospects}. Additionally, the
sensitivity of interferometers will be increased to capture more details of GW.
Having instruments with higher sensitivity, GW models of equal or higher
accuracy than observations should be used to extract GW information. Otherwise,
the extracted information would be affected more by GW models instead of
interferometer sensitivity, resulting in a bottleneck in GW analyses. Although
GW models are accurate enough for current analyses, the accuracy of current
models will no longer suffice for future data analyses
\citep{purrer2020gravitational}. Hence, it is necessary for us to develop and
improve GW models.

% Waveform families and slight motivation of why we take Phenom model
Currently, three families of GW models are commonly used. They are the
effective-one-body (EOB) \citep{ossokine2020multipolar, cotesta2020frequency, taracchini2014effective}, Numerical Relativity
(NR) surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2} and phenomenological (Phenom) models
\citep{husa2016frequency, khan2016frequency, garcia2020multimode, pratten2021computationally}. 
EOB models are constructed by mapping two masses onto an
effective body under an effective metric; NR surrogate models construct
waveforms using combinations of NR waveforms; Phenom models are formulated using
specific ansatz and inspiral approximations. While EOB and NR surrogate models
give better waveform approximants, Phenom waveforms can be produced much faster,
hence it is used mostly in data analysis tasks that requires many waveform
generations. This advantage scales up in data analysis tasks such as matched
filtering and parameter estimation, where many waveforms are required in each
run. This motivates us to improve upon the current framework of Phenom models,
thus can retain the advantage of fast waveform generation while improving the
model's accuracy.

>>>>

% Brief intro of waveform models

Many data analysis tasks in today's gravitational wave (GW) astrophysics, such
as match filtering search and parameter estimation, rely on accurate waveform
models. Because using numerical relativity (NR) waveforms in these tasks is
prohibitively expensive, the community has constructed waveform approximants
that can be evaluated much faster. There are three families of commonly used GW
approximants, which are effective-one-body (EOB) \citep{ossokine2020multipolar,
cotesta2020frequency, taracchini2014effective}, Numerical Relativity (NR)
surrogate \citep{islam2022surrogate, varma2019surrogate, varma2019surrogate2}
and phenomenological (Phenom) models \citep{husa2016frequency,
khan2016frequency, garcia2020multimode, pratten2021computationally}. While the
detail of construction of these models are different, they all have a set of
internal parameters that can be calibrated to NR waveforms. The quality of a
waveform model is determined by the ansatz used and the accuracy of the
calibrated parameters.

% Why there is a need

The LIGO-VIRGO-KAGRA (LVK) collaboration
(\citep{LIGOScientific:2014pky,LIGOScientific:2021usb,LIGOScientific:2021djp,
VIRGO:2014yos,KAGRA:2020tym}.) has recently started their fourth observational
run on May 26, 2023. This new run is expected to double the rate of current binary black
hole (BBH) observations \citep{abbott2020prospects}. Additionally, the
sensitivity of interferometers will be increased to capture more details of GW.
Having instruments with higher sensitivity, GW models of equal or higher
accuracy than observations should be used to extract GW information. Otherwise,
the extracted information would be affected more by GW models instead of
interferometer sensitivity, resulting in a bottleneck in GW analyses. Although
GW models are accurate enough for current analyses, the accuracy of current
models will no longer suffice for future data analyses
\citep{purrer2020gravitational}. Hence, it is necessary for us to develop and
improve GW models.

% AD waveforms
